State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-04-15 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
sync repositories with everyone

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% \documentclass[twocolumn,11pt,a4paper]{article}		%<-use this for checking equation length
\documentclass[12pt]{iopart}		%<-comment out this for checking equations (author list too)
\usepackage{graphicx, algorithm, algorithmicx, algpseudocode,epsfig}
\usepackage{color}  
\usepackage{iopams,setstack}                  %<-comment out this for checking equations
% \usepackage{anysize}							%<-use this for checking equation length
% \marginsize{1.7cm}{1.5cm}{0.3cm}{0.8cm}			%<-use this for checking equation length
% \setlength{\columnsep}{0.5cm}					%<-use this for checking equation length
% \usepackage{amsmath,amssymb,amsfonts} 			%<-use this for checking equation length
% \newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
\newcommand{\todo}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}

@@

\end{algorithmic}
\end{algorithm}
\section{Results}
The neural field model derived in section~\ref{NeuralModelSection} was used to the generate data for state and parameter estimation. All parameters for the model are given in table~\ref{}. An example of the simulated field can be seen in figure~\ref{}~A. The Figure demonstrates the experimental design where the sensors were placed in the centre of the field (full width at half maximum). This design allowed for uncertainty around the edges of the sensor array, which would be typical from physical data, and alleviated problems associated with boundary conditions. Figure~\ref{}~B. shows example data from the sensors of the model, and figure~\ref{}~C. shows the power spectral density of on of the simulated time-series. The PSD shows the typical $1/f$ characteristics of intracranial EEG.
% \begin{figure}
% 	\centering
% 		\includegraphics[scale=1]{file.eps}
% 	\caption{Example of experimental design and model output. \textbf{A}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the width ($1/2$ height of Gaussian) are shown by the circles. \textbf{B} Example of data generated by the model through the observations. \textbf{C}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.}
% 	\label{fig:ExampleFieldAndSensorsDataAndPSD}
% \end{figure}

\subsection{Spatial Frequency Analysis}
The spatial frequency of the observations is shown in figure~\ref{}. The peak power of the fft of the observation was approximately $30$ dB. The cut-off frequency was taken to correspond to half the peak amplitude which was approximately $15$~dB. This provided a conservative The oversampling parameter $\rho=2$. From equation~\ref{}, this gives a minimum distance between basis functions of $2.7$ to satisfy Shannon's sampling criterion. \todo{maybe from Parham's derivation... From equation~\ref{} the width was calculated to be... or...}. The width of the basis functions was chosen such that they had sufficient overlap to represent the dynamic field.
Following this, the distance between basis function, $\Delta_b$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 3.6.
\subsection{State and Parameter Estimates}
To demonstrate the performance of the parameter and state estimation, $200$ realisations of the data were generated. Each realisation consisted of 300 ms of data and the estimation was applied to the final 200 ms, allowing the model to stabalise. The initial state and parameters were unknown to the estimator. The state estimates were evaluated by comparing the field reconstruction to the true field using the mean RMSE and standard deviation of the RMSE. The distribution of the state estimates are shown in figure~\ref{}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% \documentclass[twocolumn,11pt,a4paper]{article}		%<-use this for checking equation length
\documentclass[12pt]{iopart}		%<-comment out this for checking equations (author list too)
\usepackage{iopams,setstack}                  %<-comment out this for checking equations
\usepackage{graphicx, algorithm, algorithmicx, algpseudocode,epsfig}
\usepackage{color}  
% \usepackage{anysize}							%<-use this for checking equation length
% \marginsize{1.7cm}{1.5cm}{0.3cm}{0.8cm}			%<-use this for checking equation length
% \setlength{\columnsep}{0.5cm}					%<-use this for checking equation length
% \usepackage{amsmath,amssymb,amsfonts} 			%<-use this for checking equation length
\newcommand{\todo}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}

@@

\end{algorithmic}
\end{algorithm}
\section{Results}
The neural field model derived in section~\ref{NeuralModelSection} was used to the generate data for state and parameter estimation. All parameters for the model are given in table~\ref{}. An example of the simulated field can be seen in figure~\ref{}~A. The Figure demonstrates the experimental design where the sensors were placed in the centre of the field (full width at half maximum). This design allowed for uncertainty around the edges of the sensor array, which would be typical from physical data, and alleviated problems associated with boundary conditions. Figure~\ref{}~B. shows example data from the sensors of the model, and figure~\ref{}~C. shows the power spectral density of on of the simulated time-series. The PSD shows the typical $1/f$ characteristics of intracranial EEG.
% \begin{figure}
% 	\centering
% 		\includegraphics[scale=1]{file.eps}
% 	\caption{Example of experimental design and model output. \textbf{A}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{B} Example of data generated by the model through the observations. \textbf{C}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.}
% 	\label{fig:ExampleFieldAndSensorsDataAndPSD}
% \end{figure}

\subsection{Spatial Frequency Analysis}
The spatial frequency of the observations is shown in figure~\ref{}. The peak power of the fft of the observation was approximately $30$ dB. The cut-off frequency was taken to correspond to half the peak amplitude which was approximately $15$~dB. This provided a conservative The oversampling parameter $\rho=2$. From equation~\ref{}, this gives a minimum distance between basis functions of $2.7$ to satisfy Shannon's sampling criterion. \todo{maybe from Parham's derivation... From equation~\ref{} the width was calculated to be... or...}. The width of the basis functions was chosen such that they had sufficient overlap to represent the dynamic field.
Following this, the distance between basis function, $\Delta_b$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 3.6.
\subsection{State and Parameter Estimates}
To demonstrate the performance of the parameter and state estimation, $200$ realisations of the data were generated. Each realisation consisted of 300 ms of data and the estimation was applied to the final 200 ms, allowing the model to stabilise. The initial state and parameters were unknown to the estimator. The state estimates were evaluated by comparing the field reconstruction to the true field using the mean over time of the RMSE and standard deviation of the RMSE. The distribution of the state estimates are shown in figure~\ref{}.
